### 2023-08-17 Preliminary Analyses/Preliminary Data Analysis.Rmd
Full prelim analyses completed. For 2024 MSC proposal

---
title: "Preliminary Data Analysis"
author: "Micah E. Hirsch"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document shows the results of the preliminary data analysis for the older adult perceptual learning study. The results shown here are primarily descriptive results and data visualizations to include in a proposal document for the 2024 Motor Speech Conference. Full data analysis will be conducted at a later time.

All data processing and analysis was conducted using R Version 4.3.1

```{r, echo = F, warning = F, message = F}

# Loading in needed packages

library(tidyverse)
library(rio)
library(gt)
library(reshape2) #install.packages("reshape2")
library(ggcorrplot) #install.packages("ggcorrplot")
library(patchwork) #install.packages("patchwork")

# Loading in cleaned dataset

pl_data <- rio::import("Cleaned_Data.csv")

```

# Descriptive Statistics

## Intelligibility Data

We first started by calculating pretest and posttest intelligibility for each listener. This was done by summing the total number of correctly-identified words and the total number of target words presented across the phrases in the pretest and posttest phases. Then, intelligibility was calculated by dividng the total number of correct words by the total number of target words and multiplying by 100. Intelligibility improvement (posttest - pretest intelligibility) was then calculated for each listener. Click "show code" to see this process in more detail.

```{r, warning = F, message = F}

# Calculating Intelligibility

intel_data <- pl_data %>%
  ## Grouping by id (i.e. each listener in the study) and testing phase (pretest, posttest)
  dplyr::group_by(id, type) %>%
  ## Calculating total number of correct words (words_c), total number of target words (words_t),
  ## pretest and posttest intelligibility for each listener (intel).
  dplyr::summarize(words_c = sum(correct_words), 
                   words_t = sum(target_number),
                   intel = (words_c/words_t)*100) %>%
  ## Removing words_c and words_t since it is no longer needed
  dplyr::select(!c(words_c, words_t)) %>%
  tidyr::pivot_wider(names_from = type,
                     values_from = intel) %>%
  ## Calculating improvement scores for each listener
  dplyr::mutate(improv = posttest - pretest) %>%
  dplyr::ungroup()

# Merging the speaker and cognitive data back into the dataset. This is the main dataset we will use for analyses

## First getting relevant information from the primary dataset.
## For this analysis, we are using the age-corrected cognitive scores (variables that end in _c)
analysis_data <- pl_data %>%
  dplyr::select(c(id, speaker, list_sort_c:win_l)) %>%
  dplyr::distinct() %>%
  dplyr::right_join(intel_data, ., by = "id")

# Removing unneeded items from the environment

rm(intel_data)

```

### Descriptive Intelligibility Data Across Listeners

Across listeners, the mean and standard deviation values for pretest intelligibility, posttest intelligibility, and intelligibility difference is below.

```{r, warning = F, message = F}

intel_overall <- analysis_data %>%
    dplyr::select(id:speaker) %>%
    tidyr::pivot_longer(cols = posttest:improv,
                        names_to = "phase", 
                        values_to = "intel") %>%
    dplyr::mutate(phase = as.factor(phase),
                  phase = fct_relevel(phase, "pretest", "posttest", "improv"),
                  phase = recode(phase, pretest = "Pretest", posttest = "Posttest",
                                 improv = "Difference")) %>%
    dplyr::group_by(phase) %>%
    dplyr::summarize(Mean = mean(intel), SD = sd(intel)) %>%
    dplyr::ungroup() %>%
    gt::gt() %>%
    cols_label(phase = " ")

intel_overall

# Saving Table

intel_overall %>%
  gt::gtsave("Overall_Intel.html", path = "Tables")

```

### Intelligibility Improvement by Training Group

Listeners in this study were either assigned to perceptual training conditions for a male speaker with cerebellar ataxia (AM1) or a male speaker with Parkinson's disease (PDM10). We also calculated the mean and standard deviation in pretest intelligibility, posttest intelligibility, and intelligibility difference for each training group.

```{r, warning = F, message = F}

by_speaker <- analysis_data %>%
    dplyr::select(id:speaker) %>%
    tidyr::pivot_longer(cols = posttest:improv,
                        names_to = "phase", 
                        values_to = "intel") %>%
    dplyr::mutate(phase = as.factor(phase),
                  phase = fct_relevel(phase, "pretest", "posttest", "improv"),
                  phase = recode(phase, pretest = "Pretest", posttest = "Posttest",
                                 improv = "Difference")) %>%
    dplyr::group_by(speaker, phase) %>%
    dplyr::summarize(Mean = mean(intel), SD = sd(intel)) %>%
    dplyr::ungroup() %>%
    tidyr::pivot_wider(names_from = speaker,
                       values_from = c(Mean:SD)) %>%
    dplyr::relocate(SD_AM1, .before = Mean_PDM10) %>%
    gt::gt() %>%
    fmt_number(columns = "Mean_AM1":"SD_PDM10", decimals = 2) %>%
    tab_spanner(label = "AM1",
                columns = c(Mean_AM1, SD_AM1)) %>%
    tab_spanner(label = "PDM10",
                columns = c(Mean_PDM10, SD_PDM10)) %>%
    cols_label(phase = " ",
               Mean_AM1 = "Mean",
               SD_AM1 = "SD",
               Mean_PDM10 = "Mean",
               SD_PDM10 = "SD")

by_speaker

# Saving the Table

by_speaker %>%
  gt::gtsave("Intel_by_speaker.html", path = "Tables")

# Removing unneeded items

rm(by_speaker, intel_overall)

```

### Pretest vs Posttest Plot

A visual showing difference in pretest and posttest intelligibility scores for all listeners. The mean change in intelligibility for each speaker group is noted by the darker bolder lines. Note the variability in the individual listener responses.

```{r, warning = F, message = F}

# Creating a Custom Color Palette
my_pal <- c("#160F29", "#246A73", "#368F8B", "#F3DFC1", "#DDBEA8")

pre_post <- analysis_data %>%
  dplyr::select(c(id:pretest, speaker)) %>%
  tidyr::pivot_longer(cols = posttest:pretest,
                      names_to = "phase",
                      values_to = "intel") %>%
  dplyr::mutate(phase = as.factor(phase),
                phase = fct_relevel(phase, "pretest", "posttest")) %>%
  ggplot() +
  aes(x = phase,
      y = intel,
      group = id,
      color = speaker) +
  geom_line(alpha = 0.3) +
  stat_summary(aes(group = speaker),
               fun.y = mean, geom = "line", linewidth = 2) +
  facet_wrap("speaker") +
  scale_color_manual(values = my_pal) +
  labs(x = "Phase", y = "Intelligibility") +
  theme_bw() +
  theme(legend.position = "none", aspect.ratio = 1)

pre_post

# Save the plot

ggsave("Plots/Pre_Post.png", plot = last_plot())

```

## Cognitive Subtests

In this section, we will now look at the listeners' scores on the subtests from the NIH Toolbox Cognitive Battery. The assessments that were administered to the participants are listed below. In this analysis, we are using the age-corrected standard scores.

-   List Sort Working Memory Test (list_sort_c)

-   Flanker Inhibitory Control and Attention Test (flanker_c)

-   Dimensional Change Card Sort Test (card_sort_c)

    -   We also refer to this as cognitive flexibility

-   Picture Vocabulary Test (vocab_c)

-   Pattern Comparison Processing Speed Test (pattern_c)

-   Picture Sequence Memory Test (pic_seq_c)

The dataset also contains the words-in--noise thresholds for the participants' right (win_r) and left (win_l). We determined the participant's best threshold (e.g. the ear with the lower threshold score) and created a new variable called best_win. To see the full data processing steps, click "show code".

```{r}

# Determining ear with the lowest words-in-noise threshold (best_win)

analysis_data <- analysis_data %>%
  dplyr::mutate(best_win = ifelse(win_r < win_l, win_r, win_l)) %>%
  dplyr::select(!c(win_r, win_l, fluid_cog_c))

```

### Overall Cognitive Scores

The descriptive statistics for performance on each of the cognitive subtests are

```{r, warning = F, message = F}

cog_des <- analysis_data %>%
  dplyr::select(!c(posttest:speaker)) %>%
  tidyr::pivot_longer(cols = list_sort_c:best_win,
                      names_to = "cog",
                      values_to = "score") %>%
  dplyr::group_by(cog) %>%
  dplyr::summarize(Mean = mean(score, na.rm = T), SD = sd(score, na.rm = T)) %>%
  gt::gt() %>%
  fmt_number(columns = "Mean":"SD", decimals = 2) %>%
  cols_label(cog = " ")

cog_des

cog_des %>%
  gt::gtsave("Overall_cog.html", path = "Tables")

```

### Cognitive Subtest Scores per Training Group

We are now comparing the mean and standard deviation in these cognitive scores for listeners who were trained with AM1 vs PDM10.

```{r, warning = F, message = F}

cog_by_speaker <- analysis_data %>%
  dplyr::select(!c(posttest:improv)) %>%
  tidyr::pivot_longer(cols = list_sort_c:best_win,
                      names_to = "cog",
                      values_to = "score") %>%
  dplyr::group_by(cog, speaker) %>%
  dplyr::summarize(Mean = mean(score, na.rm = T), SD = sd(score, na.rm = T)) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(names_from = speaker,
                     values_from = c(Mean:SD)) %>%
  dplyr::relocate(SD_AM1, .before = Mean_PDM10) %>%
  gt::gt() %>%
  fmt_number(columns = "Mean_AM1":"SD_PDM10", decimals = 2) %>%
  tab_spanner(label = "AM1",
              columns = c(Mean_AM1, SD_AM1)) %>%
  tab_spanner(label = "PDM10",
              columns = c(Mean_PDM10, SD_PDM10)) %>%
  cols_label(cog = " ",
             Mean_AM1 = "Mean",
             SD_AM1 = "SD",
             Mean_PDM10 = "Mean",
             SD_PDM10 = "SD")

cog_by_speaker

cog_by_speaker %>%
  gt::gtsave("cog_by_speaker.html", path = "Tables")

```

# Correlations

In this study, we are particularly interested in the relationships between these cognitive subtests and initial intelligibility (pretest) and intelligibility difference between pretest and posttest (improv). In this preliminary analysis, we first started by calculating the correlations between these measures for 1) listeners trained with AM1, and 2) listeners trained with PDM10.

## Correlation Matrix for Listeners assigned to AM1

```{r, warning = F, message = F}

AM1_corr <- analysis_data %>%
  dplyr::filter(speaker == "AM1") %>%
  dplyr::select(c(pretest:improv, list_sort_c:best_win)) %>%
  dplyr::rename("Initial Intelligibility" = pretest,
                "Intelligibility Improvement" = improv,
                "Working Memory" = list_sort_c,
                "Inhibitory Control" = flanker_c,
                "Cognitive Flexibility" = card_sort_c,
                "Receptive Vocabulary" = vocab_c,
                "Processing Speed" = pattern_c,
                "Speech in Noise Threshold" = best_win,
                "Picture Sequence Memory" = pic_seq_c)

AM1_corr <- round(cor(AM1_corr, use = "pairwise.complete.obs"), 2)

get_lower_tri<-function(AM1_corr){
    AM1_corr[upper.tri(AM1_corr)]<- NA
    return(AM1_corr)
  }

AM1_lower_cor <- get_lower_tri(AM1_corr)

AM1_lower_cor <- melt(AM1_lower_cor, na.rm = T)

AM1 <- AM1_lower_cor %>%
  ggplot() +
  aes(x = Var1,
      y = Var2,
      fill = value) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = my_pal[1], high = my_pal[1], mid = "white",
                      midpoint = 0, limit = c(-1,1), space = "Lab",
                      name = "Pearson\nCorrelation") +
  geom_text(aes(Var1, Var2, label = value), color = "black", size = 3) +
  labs(x = " ",
       y = " ") +
  ggtitle('AM1') +
  theme_bw() +
  theme(axis.text.x = element_text(color = "black", angle = 45, hjust = 1),
        aspect.ratio = 1)

AM1

ggsave("Plots/AM1_corr.png", plot = last_plot())

```

## Correlation Martix for Listeners Assigned to PDM10

```{r, warning = F, message = F}

PDM10_corr <- analysis_data %>%
  dplyr::filter(speaker == "PDM10") %>%
  dplyr::select(c(pretest:improv, list_sort_c:best_win)) %>%
  dplyr::rename("Initial Intelligibility" = pretest,
                "Intelligibility Improvement" = improv,
                "Working Memory" = list_sort_c,
                "Inhibitory Control" = flanker_c,
                "Cognitive Flexibility" = card_sort_c,
                "Receptive Vocabulary" = vocab_c,
                "Processing Speed" = pattern_c,
                "Speech in Noise Threshold" = best_win,
                "Picture Sequence Memory" = pic_seq_c)

PDM10_corr <- round(cor(PDM10_corr, use = "pairwise.complete.obs"), 2)

get_lower_tri<-function(PDM10_corr){
    PDM10_corr[upper.tri(PDM10_corr)]<- NA
    return(PDM10_corr)
  }

PDM10_lower_cor <- get_lower_tri(PDM10_corr)

PDM10_lower_cor <- melt(PDM10_lower_cor, na.rm = T)

PDM10 <- PDM10_lower_cor %>%
  ggplot() +
  aes(x = Var1,
      y = Var2,
      fill = value) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = my_pal[2], high = my_pal[2], mid = "white",
                      midpoint = 0, limit = c(-1,1), space = "Lab",
                      name = "Pearson\nCorrelation") +
  geom_text(aes(Var1, Var2, label = value), color = "black", size = 3) +
  labs(x = " ",
       y = " ") +
  ggtitle('PDM10') +
  theme_bw() +
  theme(axis.text.x = element_text(color = "black", angle = 45, hjust = 1),
        aspect.ratio = 1)

PDM10

ggsave("Plots/PDM10_corr.png", plot = last_plot())

```

```{r, echo = F, eval = F}

figure_1 <- (AM1 | PDM10) /
  pre_post

figure_1

ggsave("Plots/Full_figure.png", plot = last_plot(), width = 15, height = 15, units = "in")

```