### 2024-01-04 Manuscript Analyses/manuscript_analysis.Rmd
Knit draft rmd report

---
title: Cognitive and Auditory Predictors of Dysarthric Speech Understanding in Older
  Adult Listeners
author: "Lansford, Hirsch, Barrett, & Borrie"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document contains the analysis code and output for the manuscript entitled "Cognitive and Auditory Predictors of Dysarthric Speech Understanding in Older Adult Listeners". These results will also be presented at the Biannual Motor Speech Conference held in San Diego, CA from February 21-24, 2024.

The data preprocessing script can be found in the Manuscript Analyses subdirectory in this repo.

All data processing and analyses were conducted using R Version 4.3.2.

```{r, echo = F, warning = F, message = F}

# Loading in needed packages
library(tidyverse) #install.packages("tidyverse")
library(rio) #install.packages("rio")
library(gt) #install.packages("gt")
library(gtsummary) #install.packages("gtsummary")
library(reshape2) #install.packages("reshape2")
library(ggcorrplot) #install.packages("ggcorrplot")
library(patchwork) #install.packages("patchwork")
library(janitor) #install.packages("janitor")
library(fs) #install.packages("fs")
library(furniture) #install.packages("furniture")
library(caret) #install.packages("caret")
library(effectsize) #install.packages("effectsize")
library(parameters) #install.packages("parameters")
library(performance) #install.packages("performance")
library(bayestestR) #install.packages("bayestestR)
library(rstanarm) #install.packages("rstanarm")
library(afex) #install.packages("afex")
library(glmnet) #install.packages("glmnet")

# Load in datasets

## Listener Demographic Information
demo <- rio::import("participant_demo.csv")

## Listener Intelligibility and Cognition Scores
data <- rio::import("cleaned_data.csv")

```

First, pretest and posttest intelligibility was calculated as percent words correct (pwc) for each listener.

```{r, warning = F, message = F}

# Calculating pretest and posttest intelligibility for the listeners
pl_int <- data %>%
  dplyr::group_by(id, speaker, type) %>%
  dplyr::mutate(pwc = (sum(correct_words)/sum(target_number))*100,
                best_win = ifelse(win_r < win_l, win_r, win_l)) %>%
  dplyr::summarize(pwc = mean(pwc, na.rm = T),
                   across(c(contains("_c"), contains("_u"), contains("win")),
                          ~first(.x)
                          )) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(type = as.factor(type),
                type = fct_relevel(type, "pretest", "posttest"))

# Subsetting age from demo df

age <- demo %>%
  select(id, age)

# Merging listener age into pl_int

pl_int <- dplyr::left_join(age, pl_int, by = "id")

cor_df <- pl_int %>%
  tidyr::pivot_wider(names_from = type,
                     values_from = pwc) %>%
  dplyr::mutate(improvement = posttest - pretest) %>%
  dplyr::select(id, speaker, ends_with("_c"), best_win, pretest, posttest, improvement)

```

```{r, echo = F, message = F, warning = F}

# removing unneeded objects from the environment
rm(age, data)

```

# Descriptives

## Participant Demographics

A total of 60 older adult listeners participated in the study. The overall demographic information for the participants are below.

```{r}

demo_table <- demo %>%
  dplyr::select(-id) %>%
  tbl_summary(type = list(age ~ "continuous",
                          gender ~ "categorical",
                          race ~ "categorical",
                          ethnicity ~ "categorical"),
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                              all_categorical() ~ "{n} ({p}%)"),
              digits = list(everything() ~ c(2))) %>%
  as_gt()

demo_table

demo_table %>%
  gt::gtsave("Tables/demo_table.html")

```

```{r, echo = F, warning = F, message = F}

# Removing unneeded items from the environment

rm(demo, demo_table)

```

## Cognition

In this study, listeners were assigned to one of two speakers for perceptual training: a speaker with Ataxic dysarthria (AM1) or a speaker with Parkinson's disease (PDM10). Means and standard deviation for listener age, cognition, and word in noise hearing threshold for each condition are presented below. Wilcoxon rank sum tests were performed for each of the measures. None of the cognition or hearing measures were significantly different between the two groups of listeners.

```{r, warning = F, message = F}

cog_table <- pl_int %>%
  dplyr::filter(type == "pretest") %>%
  dplyr::select(id, speaker, age, ends_with("_c"), best_win) %>%
  dplyr::select(-fluid_cog_c) %>%
  dplyr::rename(
    `Working Memory` = list_sort_c,
    `Inhibitory Control` = flanker_c,
    `Cognitive Flexibility` = card_sort_c,
    `Receptive Vocabulary` = vocab_c,
    `Processing Speed` = pattern_c,
    `Picture Sequence Memory` = pic_seq_c,
    `Words in Noise Threshold` = best_win
  ) %>%
  group_by(id, speaker) %>% 
  summarize(across(everything(), first)) %>% 
  ungroup() %>% 
  select(-id) %>% 
  tbl_summary(
    by = "speaker",
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no",
    digits = list(everything() ~ c(2))
  ) %>% 
  add_p() %>%
  as_gt()

cog_table

cog_table %>%
  gt::gtsave("Tables/means_tabel.html")

```

## Intelligibility

Pretest and posttest intelligibility, as measured by percent words correct, for each of the two speaker conditions are presented below. On average, pretest intelligibility for AM1 was lower than PDM10. However, the listener's who were trained with AM1, on average, had greater intelligibility gains at posttest compared to the listener's trained with PDM10. A repeated measures ANOVA was conducted on the pretest-posttest change for the AM1 was significant (F(1,32) = 103.69, p < .001, d = 1.01).

```{r, warning = F, message = F}

# Intelligibility Descriptives Table

speaker_intel <- pl_int %>%
  dplyr::mutate(type = as.factor(type),
                type = fct_relevel(type, "pretest", "posttest")) %>%
  dplyr::group_by(speaker, type) %>%
  dplyr::summarize(m_pwc = mean(pwc),
                   sd_pwc = sd(pwc)) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(names_from = speaker,
                     values_from = c(m_pwc:sd_pwc)) %>%
  dplyr::relocate(sd_pwc_AM1, .after = m_pwc_AM1) %>%
  gt::gt() %>%
    fmt_number(columns = "m_pwc_AM1":"sd_pwc_PDM10", decimals = 2) %>%
    tab_spanner(label = "AM1",
                columns = c(m_pwc_AM1, sd_pwc_AM1)) %>%
    tab_spanner(label = "PDM10",
                columns = c(m_pwc_PDM10, sd_pwc_PDM10)) %>%
    cols_label(type = " ",
               m_pwc_AM1 = "Mean",
               sd_pwc_AM1 = "SD",
               m_pwc_PDM10 = "Mean",
               sd_pwc_PDM10 = "SD")

speaker_intel

speaker_intel %>%
  gt::gtsave("Tables/intel_des.html")

```

### Pretest - Posttest Plot

```{r, warning = F, message = F}

pl_int %>%
  ggplot() +
  aes(x = type,
      y = pwc,
      group = id) +
  geom_line(alpha = 0.3) +
  stat_summary(aes(group = speaker),
               fun.y = mean, geom = "line", linewidth = 2) +
  facet_wrap("speaker") +
  labs(x = "Phase", y = "Intelligibility") +
  theme_classic() +
  theme(legend.position = "none", aspect.ratio = 1)

ggsave("Plots/Pre_post.png", plot = last_plot())

```

```{r}

cor_df %>% 
  dplyr::select(id, speaker, pretest, posttest) %>% 
  tidyr::pivot_longer(pretest:posttest) %>% 
  dplyr::group_by(speaker) %>% 
  dplyr::group_nest() %>% 
  dplyr::mutate(
    mod = map(data, ~afex::aov_4(value ~ 1 + (name|id), data = .x))
  ) %>% 
  pull(mod)

cor_df %>% 
  dplyr::select(id, speaker, pretest, posttest) %>% 
  tidyr::pivot_longer(pretest:posttest) %>% 
  dplyr::group_by(speaker) %>% 
  dplyr::group_nest() %>% 
  dplyr::mutate(
    mod = map(data, ~effectsize::cohens_d(value ~ name, data = .x))
  ) %>% 
  pull(mod)

```

# Results

## Correlations

The correlation matricies between all the cognitive measures, hearing threshold, initial intelligibility, and intelligibility improvement were conducted for each speaker condition.

### AM1

```{r, warning = F, message = F}

# Customize plot colors

library(LaCroixColoR) #remotes::install_github("johannesbjork/LaCroixColoR")
my_pal <- lacroix_palette("PeachPear", n = 15, type = "continuous")
my_pal <- rev(my_pal)

# AM1 Cor Plot
AM1_cor <- cor_df %>%
  dplyr::filter(speaker == "AM1") %>%
  dplyr::select(!c(id, speaker, fluid_cog_c, posttest)) %>%
  dplyr::rename(
    `Working Memory` = list_sort_c,
    `Inhibitory Control` = flanker_c,
    `Cognitive Flexibility` = card_sort_c,
    `Receptive Vocabulary` = vocab_c,
    `Processing Speed` = pattern_c,
    `Picture Sequence Memory` = pic_seq_c,
    `Words in Noise Threshold` = best_win
  ) %>%
  dplyr::rename_all(snakecase::to_title_case) %>%
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot(
    type = "lower",
    outline.col = "white",
    colors = c(my_pal[1], "white", my_pal[15]),
    lab = TRUE,
    sig.level = 0.05,
    insig = "blank",
    title = "Ataxic", 
    legend.title = "Correlation"
  )

AM1_cor

```

### PDM10

```{r, warning = F, message = F}

# PDM10 Cor Plot
PDM10_cor <- cor_df %>%
  dplyr::filter(speaker == "PDM10") %>%
  dplyr::select(!c(id, speaker, fluid_cog_c, posttest)) %>%
  dplyr::rename(
    `Working Memory` = list_sort_c,
    `Inhibitory Control` = flanker_c,
    `Cognitive Flexibility` = card_sort_c,
    `Receptive Vocabulary` = vocab_c,
    `Processing Speed` = pattern_c,
    `Picture Sequence Memory` = pic_seq_c,
    `Words in Noise Threshold` = best_win
  ) %>%
  dplyr::rename_all(snakecase::to_title_case) %>%
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot(
    type = "lower",
    outline.col = "white",
    colors = c(my_pal[1], "white", my_pal[15]),
    lab = TRUE,
    insig = "blank",
    title = "Parkinson's Disease", 
    legend.title = "Correlation"
  )

PDM10_cor

```

```{r, echo = F, warning = F, message = F}

library(patchwork)
AM1_cor + PDM10_cor + plot_layout(guides = "collect") & theme(legend.position = "bottom")
ggsave("Plots/cor_fig.png", width = 20, height = 6, dpi = 300, units = "in")

rm(AM1_cor, PDM10_cor, cog_table, speaker_intel)

```

## Models of Initial Intelligibility

```{r, warning = F, message = F}

# Functions used in this code

## Plot importance function
plot_importance = function(mods, labels){
  map2(mods, labels, ~var_imp(.x, .y)) %>% 
    reduce(full_join, by = "var") %>% 
    ungroup() %>% 
    pivot_longer(-var) %>% 
    group_by(var) %>% 
    mutate(avg_imp = mean(value, na.rm=TRUE)) %>% 
    ungroup() %>% 
    arrange(avg_imp) %>% 
    mutate(var = fct_inorder(var))
}

## Function for determining variable importance
var_imp <- function(mod, label) {
  varImp(mod)$importance %>%
    rownames_to_column(var = "Variable") %>%
    mutate(Overall = Overall/max(Overall)) %>%
    rename(!!label := Overall, 
          var = Variable)
}

## Function for model training
training <- function(data, method, int) {
  form = if (isTRUE(int)) {
    d2 = model.matrix(pwc ~ . * ., data = data)[, -1] %>%
      as_tibble() %>%
      janitor::clean_names()
    d2$pwc = data$pwc
  } else {
    d2 = data
  }
    train(
      pwc ~ ., 
      data = d2, 
      preProcess = c("center", "scale"),
      method = method, 
      trControl = caret::trainControl(
        method = "repeatedcv",
        number = 5,
        repeats = 10
    )
  )
}

```

### AM1 Models

```{r}

set.seed(7036)

mod_fun = function(speaker, type){
  subsetted = pl_int %>% 
    filter(type == {{type}}) %>% 
    filter(speaker == {{speaker}}) %>% 
    select(pwc, ends_with("_c"), best_win) %>% 
    drop_na(pwc, ends_with("_c"), best_win) %>% 
    select(-contains("fluid"))

  mod.1 = training(subsetted, "glmnet", int = FALSE)
  mod.2 = training(subsetted, "glmnet", int = TRUE)

  return(list(mod.1, mod.2))
}

plot_fun = function(mods, title){
  plot_importance(
    mods = mods,
    labels = c("main", "int")
  ) %>% 
    drop_na(value) %>% 
    mutate(name = factor(
      name, 
      levels = c("main", "int"),
      labels = c(
        paste0("Main Effects"),
        paste0("Interactions")
    ))) %>% 
    filter(name == "Interactions") %>% 
    mutate(var = str_replace(var, "flanker_c", "Inhibitory Control of Attention"),
           var = str_replace(var, "card_sort_c", "Cognitive Flexibility"),
           var = str_replace(var, "vocab_c", "Vocabulary Knowledge"),
           var = str_replace(var, "pattern_c", "Processing Speed"),
           var = str_replace(var, "list_sort_c", "Working Memory"),
           var = str_replace(var, "pic_seq_c", "Picture Sequence Memory"),
           var = str_replace(var, "best_win", "Words in Noise Threshold"),
           var = str_replace(var, "_", " x ")) %>% 
    arrange(value) %>% 
    mutate(var = fct_inorder(var)) %>% 
    ggplot(aes(x = value, y = var, color = value)) +
    geom_segment(aes(xend = 0, yend = var)) +
    geom_point(size = 2) +
    facet_wrap(~name, scales = "free") +
    labs(x = "Relative Importance",
         y = "",
         title = {{title}}) +
    scale_color_gradientn(colors = my_pal) +
    theme_bw() +
    theme(plot.title.position = "plot")
}

mods1 = mod_fun("AM1", "pretest")
p1 = plot_fun(mods1, "a) Ataxic")
p1

```

```{r}

in_and_out <- function(mods, speaker) {
  out_sample_r2 = map_dbl(mods, ~max(.x$results$Rsquared))
  out_sample_rmse = map_dbl(mods, ~min(.x$results$RMSE)) 
  tibble(
    speaker = speaker,
    mods = c("Main Effects", "Interactions"),
    out_sample_r2,
    out_sample_rmse
  )
}

AM1_initial_mod <- in_and_out(mods1, "Ataxic")

AM1_initial_mod %>%
  gt::gt() %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE")

```

```{r}

t_for_diff <- function(mods){
  samps = resamples(list(
    main = mods[[1]],
    int = mods[[2]]
  ))
  summary(samps)
  diff(samps) %>%
    summary()
}

t_for_diff(mods1)

```

```{r}

effect_data <- function(mod){
  mod[[2]]$trainingData %>%
    mutate(preds = predict(mod[[2]]))
}

effect_plot <- function(mod, var, var2 = NULL, int = TRUE, num = 3) {
  dat = effect_data(mod)

  if (isFALSE(int)){
    dat %>%
      ggplot() +
      aes(x={{var}}, 
          y = .outcome) +
      geom_point() +
      geom_smooth(method = "lm")
  } else {
    dat %>%
      ggplot() +
      aes(x = {{var}},
          y= .outcome,
          group = cut_interval({{var2}}, num),
          color = cut_interval({{var2}}, num)) +
      geom_point() +
      geom_smooth(method = "lm") +
      theme(legend.position = "bottom")
  }
}

# AM1 Effect Plots

effect_plot(mods1, pic_seq_c, best_win) # Picture Sequence Memory * Words in Noise
effect_plot(mods1, best_win, pic_seq_c) # Words in Noise * Picture Sequence Memory
effect_plot(mods1, best_win, int = F) # Words in Noise
effect_plot(mods1, list_sort_c, vocab_c) # Working Memory * Receptive Vocab
effect_plot(mods1, vocab_c, list_sort_c) # Receptive Vocab * Working Memory
effect_plot(mods1, list_sort_c, pattern_c) # Working Memory * Processing Speed
effect_plot(mods1, pattern_c, list_sort_c) # Processing Speed * Working Memory

```

### PDM10 Models

```{r}

mods2 = mod_fun("PDM10", "pretest")
p2 = plot_fun(mods2, "b) PD")
p2

```

```{r}

PDM10_initial_mod <- in_and_out(mods2, "PD")

PDM10_initial_mod %>%
  gt::gt() %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE")

```

```{r}

# PDM10 Effect Plots

effect_plot(mods2, vocab_c, pattern_c) # Receptive Vocab * Processing Speed
effect_plot(mods2, pattern_c, vocab_c) # Processing Speed * Receptive Vocab
effect_plot(mods2, best_win, int = F) # Words in Noise
effect_plot(mods2, flanker_c, best_win) # Inhibitory Control * Words in Noise
effect_plot(mods2, best_win, flanker_c) # Words in Noise * Inhibitory Control
effect_plot(mods2, list_sort_c, pattern_c) # Working Memory * Processing Speed
effect_plot(mods2, pattern_c, list_sort_c) # Processing Speech * Working Memory
effect_plot(mods2, vocab_c, pic_seq_c) # Receptive Vocab * Picture Sequence Memory
effect_plot(mods2, pic_seq_c, vocab_c) # Picture Sequence Memory * Receptive Vocab
effect_plot(mods2, card_sort_c, vocab_c) # Cognitive Flexibility * Receptive Vocab
effect_plot(mods2, vocab_c, card_sort_c) # Receptive Vocab * Cognitive Flexibility

```

```{r, echo = F, warning = F, message = F}

# Mergining initial model summaries together
initial_mods <- rbind(AM1_initial_mod, PDM10_initial_mod)

# Combining initial model figures and saving
p1 / p2
ggsave("Plots/Initial_mods.png", plot = last_plot(), width = 8, height = 10, units = "in")

# Removing unneeded items from Environment
rm(p1, p2, mods1, mods2, AM1_initial_mod, PDM10_initial_mod)

```

## Models of Cognition and Intelligibility Change

### AM1 Models

```{r}

mod_fun2 <- function(speaker, title) {
  subsetted = cor_df %>%
    dplyr::select(-c(fluid_cog_c, improvement)) %>%
    dplyr::mutate(pwc_diff = resid(lm(posttest ~ pretest))) %>%
    dplyr::select(pwc_diff, ends_with("_c"), speaker) %>%
    tidyr::drop_na(pwc_diff, ends_with("_c"), speaker) %>%
    dplyr::filter(speaker == {{speaker}}) %>%
    dplyr::select(-speaker) %>%
    rename(pwc = pwc_diff)

  mod.1 = training(subsetted, "glmnet", int = F)
  mod.2 = training(subsetted, "glmnet", int = T)

  return(list(mod.1, mod.2))

}

mods12 = mod_fun2("AM1")
p12 = plot_fun(mods12, "a) Ataxic")
p12

```

```{r}

AM1_change_mod <- in_and_out(mods12, "Ataxic")

AM1_change_mod %>%
  gt::gt() %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE")

```

```{r}

t_for_diff(mods12)

```

```{r}

## AM1 Effect Plots

effect_plot(mods12, list_sort_c, vocab_c) # Working Memory * Receptive Vocab
effect_plot(mods12, vocab_c, list_sort_c) # Receptive Vocab * Working Memory
effect_plot(mods12, list_sort_c, int = F) # Working Memory
effect_plot(mods12, pattern_c, int = F) # Processing Speed
effect_plot(mods12, list_sort_c, flanker_c) # Working Memory * Inhibitory Control
effect_plot(mods12, flanker_c, list_sort_c) # Inhibitory Control * Working Memory

```

### PDM10 Models

```{r}

mods22 = mod_fun2("PDM10")
p22 = plot_fun(mods22, "b) PD")
p22

```

```{r}

PDM10_change_mod <- in_and_out(mods22, "PD")

PDM10_change_mod %>%
  gt::gt() %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE")

```

```{r}

t_for_diff(mods22)

```

```{r}

## PDM10 Effect Plots

effect_plot(mods22, vocab_c, pic_seq_c) # Receptive Vocab * Picture Sequence Memory
effect_plot(mods22, pic_seq_c, vocab_c) # Picture Sequence Memory * Receptive Vocab
effect_plot(mods22, vocab_c, int = F) # Receptive Vocab
effect_plot(mods22, list_sort_c, flanker_c) # Working Memory * Inhibitory Control
effect_plot(mods22, flanker_c, list_sort_c) # Inhibitory Control * Working Memory

```

```{r, echo = F, warning = F, message = F}

# Merging Model summaries together
change_mods <- rbind(AM1_change_mod, PDM10_change_mod)

# Combining and saving change model figures
p12 / p22
ggsave("Plots/Change_mods.png", plot = last_plot(), width = 8, height = 10, units = "in")

# Removing unneeded items from environment
rm(p12, p22, AM1_change_mod, PDM10_change_mod, mods12, mods22)

```

```{r, echo = F, warning = F, message = F}

# Creating Overall Model Table

initial_mods <- initial_mods %>%
  dplyr::mutate(outcome = "Initial Intelligibility")

change_mods <- change_mods %>%
  dplyr::mutate(outcome = "Intelligbility Change")

mods_table <- rbind(initial_mods, change_mods)

mods_table %>%
  gt::gt(groupname_col = "outcome") %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE") %>%
  gt::gtsave("Tables/mods_table.html")

```