### 2024-01-16 Manuscript Analyses/manuscript_analysis.Rmd
reran analyses

---
title: Cognitive and Auditory Predictors of Dysarthric Speech Understanding in Older
  Adult Listeners
author: "Lansford, Hirsch, Barrett, & Borrie"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

This document contains the analysis code and output for the manuscript entitled "Cognitive and Auditory Predictors of Dysarthric Speech Understanding in Older Adult Listeners". These results will also be presented at the Biannual Motor Speech Conference held in San Diego, CA from February 21-24, 2024.

The data preprocessing script can be found in the Manuscript Analyses subdirectory in this repo.

All data processing and analyses were conducted using R Version 4.3.2.

```{r, echo = F, warning = F, message = F}

# Loading in needed packages
library(tidyverse) #install.packages("tidyverse")
library(rio) #install.packages("rio")
library(gt) #install.packages("gt")
library(gtsummary) #install.packages("gtsummary")
library(reshape2) #install.packages("reshape2")
library(ggcorrplot) #install.packages("ggcorrplot")
library(patchwork) #install.packages("patchwork")
library(janitor) #install.packages("janitor")
library(fs) #install.packages("fs")
library(furniture) #install.packages("furniture")
library(caret) #install.packages("caret")
library(effectsize) #install.packages("effectsize")
library(parameters) #install.packages("parameters")
library(performance) #install.packages("performance")
library(bayestestR) #install.packages("bayestestR)
library(rstanarm) #install.packages("rstanarm")
library(afex) #install.packages("afex")
library(glmnet) #install.packages("glmnet")

# Load in datasets

## Listener Demographic Information
demo <- rio::import("participant_demo.csv")

## Listener Intelligibility and Cognition Scores
data <- rio::import("cleaned_data.csv")

```

First, pretest and posttest intelligibility was calculated as a percent words correct (pwc) score for each listener. Pretest PWC and PWC change (posttest-pretest) will be our outcome variables for the analyses.

```{r, warning = F, message = F}

# Removing Participant 323 from the dataset (missing demographic and cognitive data)

data <- data %>%
  dplyr::filter(id != 323)

demo <- demo %>%
  dplyr::filter(id != 323)

# Calculating pretest and posttest intelligibility for the listeners
pl_int <- data %>%
  dplyr::group_by(id, speaker, type) %>%
  dplyr::mutate(pwc = (sum(correct_words)/sum(target_number))*100,
                best_win = ifelse(win_r < win_l, win_r, win_l)) %>%
  dplyr::summarize(pwc = mean(pwc, na.rm = T),
                   across(c(contains("_c"), contains("_u"), contains("win")),
                          ~first(.x)
                          )) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(type = as.factor(type),
                type = fct_relevel(type, "pretest", "posttest"))

# Subsetting age from demo df

age <- demo %>%
  select(id, age)

# Merging listener age into pl_int

pl_int <- dplyr::left_join(age, pl_int, by = "id")

cor_df <- pl_int %>%
  tidyr::pivot_wider(names_from = type,
                     values_from = pwc) %>%
  dplyr::mutate(improvement = posttest - pretest) %>%
  dplyr::select(id, speaker, ends_with("_c"), best_win, pretest, posttest, improvement)

```

```{r, echo = F, message = F, warning = F}

# removing unneeded objects from the environment
rm(age, data)

```

# Descriptives

## Participant Demographics

A total of 60 older adult listeners participated in the study. The overall demographic information for the participants are in the table below.

```{r}

demo_table <- demo %>%
  dplyr::select(-id) %>%
  tbl_summary(type = list(age ~ "continuous",
                          gender ~ "categorical",
                          race ~ "categorical",
                          ethnicity ~ "categorical"),
              statistic = list(all_continuous() ~ "{mean} ({sd})",
                              all_categorical() ~ "{n} ({p}%)"),
              digits = list(everything() ~ c(2))) %>%
  as_gt()

demo_table

demo_table %>%
  gt::gtsave("Tables/demo_table.html")

```

```{r, echo = F, warning = F, message = F}

# Removing unneeded items from the environment

rm(demo, demo_table)

```

## Cognition

In this study, listeners were assigned to one of two speakers for perceptual training: a speaker with Ataxic dysarthria (AM1) or a speaker with Parkinson's disease (PDM10). Means and standard deviation for listener age, cognition, and word in noise hearing threshold for each condition are presented below. Wilcoxon rank sum tests were performed for each of the measures. None of the cognition or hearing measures were significantly different between the two groups of listeners.

```{r, warning = F, message = F}

cog_table <- pl_int %>%
  dplyr::filter(type == "pretest") %>%
  dplyr::select(id, speaker, age, ends_with("_c"), best_win) %>%
  dplyr::select(-c(fluid_cog_c, pic_seq_c)) %>%
  dplyr::rename(
    `Working Memory` = list_sort_c,
    `Inhibitory Control` = flanker_c,
    `Cognitive Flexibility` = card_sort_c,
    `Receptive Vocabulary` = vocab_c,
    `Processing Speed` = pattern_c,
    `Words in Noise Threshold` = best_win
  ) %>%
  group_by(id, speaker) %>% 
  summarize(across(everything(), first)) %>% 
  ungroup() %>% 
  select(-id) %>% 
  tbl_summary(
    by = "speaker",
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no",
    digits = list(everything() ~ c(2))
  ) %>% 
  add_p() %>%
  as_gt()

cog_table

cog_table %>%
  gt::gtsave("Tables/means_tabel.html")

```

## Intelligibility

Pretest and posttest intelligibility, as measured by percent words correct, for each of the two speaker conditions are presented below. On average, pretest intelligibility for AM1 was lower than PDM10. However, the listener's who were trained with AM1, on average, had greater intelligibility gains at posttest compared to the listener's trained with PDM10. A repeated measures ANOVA was conducted on the pretest-posttest change for the AM1 was significant (F(1,32) = 103.69, p \< .001, d = 1.01).

```{r, warning = F, message = F}

# Intelligibility Descriptives Table

speaker_intel <- cor_df %>%
  dplyr::select(speaker, pretest, posttest, improvement) %>%
  tbl_summary(
    by = speaker,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no",
    digits = list(everything() ~ c(2)),
    label = list(pretest ~ "Pretest Inelligibility",
                   posttest ~ "Posttest Intelligibility",
                   improvement ~ "Intelligibility Change")) %>%
  as_gt()

speaker_intel

speaker_intel %>%
  gt::gtsave("Tables/intel_des.html")

```

### Pretest - Posttest Plot

```{r, warning = F, message = F}

# Customize colors for figures
library(LaCroixColoR) #remotes::install_github("johannesbjork/LaCroixColoR")
my_pal <- lacroix_palette("PeachPear", type = "discrete")

pl_int %>%
  ggplot() +
  aes(x = type,
      y = pwc,
      group = id,
      color = speaker) +
  geom_line(alpha = 0.3) +
  geom_point(alpha = 0.3) +
  stat_summary(aes(group = speaker),
               fun.y = mean, geom = "line", linewidth = 2) +
  facet_wrap("speaker") +
  labs(x = "Phase", y = "Intelligibility") +
  theme_classic() +
  scale_color_manual(values = c(my_pal[5], my_pal[6])) +
  theme(legend.position = "none", aspect.ratio = 1)

ggsave("Plots/Pre_post.png", plot = last_plot())

```

```{r}

cor_df %>% 
  dplyr::select(id, speaker, pretest, posttest) %>% 
  tidyr::pivot_longer(pretest:posttest) %>% 
  dplyr::group_by(speaker) %>% 
  dplyr::group_nest() %>% 
  dplyr::mutate(
    mod = map(data, ~afex::aov_4(value ~ 1 + (name|id), data = .x))
  ) %>% 
  pull(mod)

cor_df %>% 
  dplyr::select(id, speaker, pretest, posttest) %>% 
  tidyr::pivot_longer(pretest:posttest) %>% 
  dplyr::group_by(speaker) %>% 
  dplyr::group_nest() %>% 
  dplyr::mutate(
    mod = map(data, ~effectsize::cohens_d(value ~ name, data = .x))
  ) %>% 
  pull(mod)

```

# Results

## Correlations

The correlation matricies between all the cognitive measures, hearing threshold, initial intelligibility, and intelligibility improvement were conducted for each speaker condition.

### AM1

Words in noise significantly correlated with pretest intelligibility for AM1 at the alpha = 0.05 level. Listeners with higher (i.e. worse) words in noise hearing thresholds tend to have lower pretest intelligibility (i.e. lower initial intelligibility) for AM1.

```{r, warning = F, message = F}

# Customize plot colors
my_pal <- lacroix_palette("PeachPear", n = 15, type = "continuous")
my_pal <- rev(my_pal)

# AM1 Cor Plot
AM1_cor <- cor_df %>%
  dplyr::filter(speaker == "AM1") %>%
  dplyr::select(!c(id, speaker, fluid_cog_c, pic_seq_c, posttest)) %>%
  dplyr::rename(
    `Working Memory` = list_sort_c,
    `Inhibitory Control` = flanker_c,
    `Cognitive Flexibility` = card_sort_c,
    `Receptive Vocabulary` = vocab_c,
    `Processing Speed` = pattern_c,
    `Words in Noise Threshold` = best_win
  ) %>%
  dplyr::rename_all(snakecase::to_title_case) %>%
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot(
    type = "lower",
    outline.col = "white",
    colors = c(my_pal[1], "white", my_pal[15]),
    lab = TRUE,
    insig = "blank",
    title = "Ataxic", 
    legend.title = "Correlation"
  )

AM1_cor

```

The plot below shows only the significant correlations in the matrix above for AM1.

```{r, warning = F, message = F}

cor_df %>%
  dplyr::filter(speaker == "AM1") %>%
  dplyr::select(!c(id, speaker, fluid_cog_c, pic_seq_c, posttest)) %>%
  dplyr::rename(
    `Working Memory` = list_sort_c,
    `Inhibitory Control` = flanker_c,
    `Cognitive Flexibility` = card_sort_c,
    `Receptive Vocabulary` = vocab_c,
    `Processing Speed` = pattern_c,
    `Words in Noise Threshold` = best_win
  ) %>%
  dplyr::rename_all(snakecase::to_title_case) %>%
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot(
    type = "lower",
    outline.col = "white",
    colors = c(my_pal[1], "white", my_pal[15]),
    lab = TRUE,
    p.mat = cor_pmat(.),
    sig.level = 0.05,
    insig = "blank",
    title = "Ataxic", 
    legend.title = "Correlation"
  )

```

### PDM10

Like AM1, words in noise hearing threshold significantly correlated with pretest intelligibility. However, pretest intelligibility is also has a negative correlation with intelligibility improvement, indicating that listeners in this group who had higher pretest intelligibility tend to have less intelligibility gains at posttest.

```{r, warning = F, message = F}

# PDM10 Cor Plot
PDM10_cor <- cor_df %>%
  dplyr::filter(speaker == "PDM10") %>%
  dplyr::select(!c(id, speaker, fluid_cog_c, pic_seq_c, posttest)) %>%
  dplyr::rename(
    `Working Memory` = list_sort_c,
    `Inhibitory Control` = flanker_c,
    `Cognitive Flexibility` = card_sort_c,
    `Receptive Vocabulary` = vocab_c,
    `Processing Speed` = pattern_c,
    `Words in Noise Threshold` = best_win
  ) %>%
  dplyr::rename_all(snakecase::to_title_case) %>%
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot(
    type = "lower",
    outline.col = "white",
    colors = c(my_pal[1], "white", my_pal[15]),
    lab = TRUE,
    insig = "blank",
    title = "Parkinson's Disease", 
    legend.title = "Correlation"
  )

PDM10_cor

```

The plot below only shows the significant correlations for PDM10.

```{r}

cor_df %>%
  dplyr::filter(speaker == "PDM10") %>%
  dplyr::select(!c(id, speaker, fluid_cog_c, pic_seq_c, posttest)) %>%
  dplyr::rename(
    `Working Memory` = list_sort_c,
    `Inhibitory Control` = flanker_c,
    `Cognitive Flexibility` = card_sort_c,
    `Receptive Vocabulary` = vocab_c,
    `Processing Speed` = pattern_c,
    `Words in Noise Threshold` = best_win
  ) %>%
  dplyr::rename_all(snakecase::to_title_case) %>%
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot(
    type = "lower",
    outline.col = "white",
    colors = c(my_pal[1], "white", my_pal[15]),
    lab = TRUE,
    p.mat = cor_pmat(.),
    insig = "blank",
    sig.level = 0.05,
    title = "Parkinson's Disease", 
    legend.title = "Correlation"
  )

```

```{r, eval = F, warning = F, message = F}

library(patchwork)
AM1_cor + PDM10_cor + plot_layout(guides = "collect") & theme(legend.position = "bottom")
ggsave("Plots/cor_fig.png", width = 20, height = 6, dpi = 300, units = "in")

```

```{r, echo = F, warning = F, message = F}

rm(AM1_cor, PDM10_cor, cog_table, speaker_intel)

```

## Models of Initial Intelligibility

```{r, warning = F, message = F}

# Functions used in this code

## Plot importance function
plot_importance = function(mods, labels){
  map2(mods, labels, ~var_imp(.x, .y)) %>% 
    reduce(full_join, by = "var") %>% 
    ungroup() %>% 
    pivot_longer(-var) %>% 
    group_by(var) %>% 
    mutate(avg_imp = mean(value, na.rm=TRUE)) %>% 
    ungroup() %>% 
    arrange(avg_imp) %>% 
    mutate(var = fct_inorder(var))
}

## Function for determining variable importance
var_imp <- function(mod, label) {
  varImp(mod)$importance %>%
    rownames_to_column(var = "Variable") %>%
    mutate(Overall = Overall/max(Overall)) %>%
    rename(!!label := Overall, 
          var = Variable)
}

## Function for model training
training <- function(data, method, int) {
  form = if (isTRUE(int)) {
    d2 = model.matrix(pwc ~ . * ., data = data)[, -1] %>%
      as_tibble() %>%
      janitor::clean_names()
    d2$pwc = data$pwc
  } else {
    d2 = data
  }
    train(
      pwc ~ ., 
      data = d2, 
      preProcess = c("center", "scale"),
      method = method, 
      trControl = caret::trainControl(
        method = "repeatedcv",
        number = 5,
        repeats = 10
    )
  )
}

```

### AM1 Models

Two elastic net models were conducted to evaluate the predictive relationship between the cognitive measures on pretest (initial) intelligibility: A model containing only the main effects of the cognitive measures (i.e. the main effects model) and another model that contains the main effects and all two-way interactions (i.e. the interaction model). The cross-validated R-squared and RMSE values are not significantly different between these two models. However, the r-squared value for the main effects model is higher. Therefore, the main effects model is the best model for the AM1 initial intelligibility model.

The relative variable importance for the models are shown in the figure below. The most important predictor for the AM1 initial intelligibility model was words-in-noise hearing threshold, followed by the main effect of working memory.

```{r, warning = F, message = F}

set.seed(7036)

mod_fun = function(speaker, type){
  subsetted = pl_int %>% 
    filter(type == {{type}}) %>% 
    filter(speaker == {{speaker}}) %>% 
    select(pwc, ends_with("_c"), best_win) %>% 
    drop_na(pwc, ends_with("_c"), best_win) %>% 
    select(-c(fluid_cog_c, pic_seq_c))

  mod.1 = training(subsetted, "glmnet", int = FALSE)
  mod.2 = training(subsetted, "glmnet", int = TRUE)

  return(list(mod.1, mod.2))
}

plot_fun = function(mods, title){
  plot_importance(
    mods = mods,
    labels = c("main", "int")
  ) %>% 
    drop_na(value) %>% 
    mutate(name = factor(
      name, 
      levels = c("main", "int"),
      labels = c(
        paste0("Main Effects"),
        paste0("Interactions")
    ))) %>% 
    filter(name == "Interactions") %>% 
    mutate(var = str_replace(var, "flanker_c", "Inhibitory Control of Attention"),
           var = str_replace(var, "card_sort_c", "Cognitive Flexibility"),
           var = str_replace(var, "vocab_c", "Receptive Vocabulary"),
           var = str_replace(var, "pattern_c", "Processing Speed"),
           var = str_replace(var, "list_sort_c", "Working Memory"),
           var = str_replace(var, "best_win", "Words in Noise Threshold"),
           var = str_replace(var, "_", " x ")) %>% 
    arrange(value) %>% 
    mutate(var = fct_inorder(var)) %>% 
    ggplot(aes(x = value, y = var, color = value)) +
    geom_segment(aes(xend = 0, yend = var)) +
    geom_point(size = 2) +
    facet_wrap(~name, scales = "free") +
    labs(x = "Relative Importance",
         y = "",
         title = {{title}}) +
    scale_color_gradientn(colors = my_pal) +
    theme_bw() +
    theme(plot.title.position = "plot")
}

mods1 = mod_fun("AM1", "pretest")
p1 = plot_fun(mods1, "a) Ataxic")
p1

```

```{r, warning = F, message = F}

in_and_out <- function(mods, speaker) {
  out_sample_r2 = map_dbl(mods, ~max(.x$results$Rsquared))
  out_sample_rmse = map_dbl(mods, ~min(.x$results$RMSE)) 
  tibble(
    speaker = speaker,
    mods = c("Main Effects", "Interactions"),
    out_sample_r2,
    out_sample_rmse
  )
}

AM1_initial_mod <- in_and_out(mods1, "Ataxic")

AM1_initial_mod %>%
  gt::gt() %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE")

```

```{r, warning = F, message = F}

t_for_diff <- function(mods){
  samps = resamples(list(
    main = mods[[1]],
    int = mods[[2]]
  ))
  summary(samps)
  diff(samps) %>%
    summary()
}

t_for_diff(mods1)

```

The plots below are showing the effects of words in noise hearing threshold and working memory on initial intelligibility for AM1.

-   Words in Noise Hearing Threshold has a negative relationship with initial intelligibility, indicating that those with higher words in noise thresholds have lower initial intelligibility for AM1.
-   Working Memory has a positive relationship with initial intelligibility, indicating that those listeners with greater working memory capacity had higher initial intelligibility for AM1.

```{r, warning = F, message = F}

effect_data <- function(mod){
  mod[[2]]$trainingData %>%
    mutate(preds = predict(mod[[2]]))
}

effect_plot <- function(mod, var, var2 = NULL, int = TRUE, num = 3) {
  dat = effect_data(mod)

  if (isFALSE(int)){
    dat %>%
      ggplot() +
      aes(x={{var}}, 
          y = .outcome) +
      geom_point() +
      geom_smooth(method = "lm")
  } else {
    dat %>%
      ggplot() +
      aes(x = {{var}},
          y= .outcome,
          group = cut_number({{var2}}, num),
          color = cut_number({{var2}}, num)) +
      geom_point() +
      geom_smooth(method = "lm") +
      theme(legend.position = "bottom")
  }
}

# AM1 Effect Plots

effect_plot(mods1, best_win, int = F) # Words in Noise
effect_plot(mods1, list_sort_c, int = F) # Working Memory

```

### PDM10 Models

Like for AM1, two elastic net models were used to model the predictive relationship between the listener cognitive measures and pretest intelligibility for PDM10: a main effects model and an interaction model. The difference in R-square and RMSE is not significantly different between the two models. However, the R-squared value is higher for the interaction model and multiple interactions were indicated as important predictors.

The important variables for the PDM10 models is shown below. The interaction between working memory and words in noise hearing threshold is the most important predictor, followed by the main effect of words in noise hearing threshold, the interaction between working memory and receptive vocabulary, the interaction between cognitive flexibility and processing speed, the interaction between inhibitory control and words in noise hearing threshold, the interaction between receptive vocabulary and processing speed, and the interaction between working memory and cognitive flexibility.

```{r, warning = F, message = F}

mods2 = mod_fun("PDM10", "pretest")
p2 = plot_fun(mods2, "b) PD")
p2

```

```{r, warning = F, message = F}

PDM10_initial_mod <- in_and_out(mods2, "PD")

PDM10_initial_mod %>%
  gt::gt() %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE")

```

```{r, warning = F, message = F}
t_for_diff(mods2)
```

The figures below are plotting the interaction effects for the PDM10 initial intelligibility models.

-   The interaction between working memory and words in noise hearing threshold indicates that working memory did not have much of an effect for listeners with lower (i.e. better) hearing thresholds on initial intelligibility for PDM10. However, working memory help compensate for listeners with elevated hearing thresholds.
-   Like with AM1, words in noise hearing threshold has a negative relationship with initial intelligibility for PDM10. 
-   The interaction between working memory and receptive vocabulary indicates that greater receptive vocabulary knowledge helped compensate for listeners with lower working memory capacity. 
-   The interaction between cognitive flexibility and processing speech indicates that greater cognitive flexibility helped compensate for listener with lower processing speed. However, greater processing speed tends to have a negative effect on initial intelligibility for listeners with greater cognitive flexibility.
-   The interaction between inhibitory control and words in noise hearing threshold indicate that inhibitory control helped compensate for listeners with slightly elevated words in noise hearing thresholds. However, inhibitory control did not have a great effect on initial intelligibility for listeners with low or elevated hearing thresholds.
-   The interaction between receptive vocabulary and processing speed indicates that those with greater processing speed and greater vocabulary had greater initial intelligibility
-   The interaction between cognitive flexibility and working memory indicates that cognitive flexibility helps compensate perception for listeners with lower working memory capacity, but may not have much of an effect (or even a possible negative effect) for listeners with higher working memory capacity.

```{r, warning = F, message = F}

# PDM10 Effect Plots

effect_plot(mods2, list_sort_c, best_win) # Working Memory * Words in Noise Threshold
effect_plot(mods2, best_win, int = F) # Words in Noise Threshold
effect_plot(mods2, vocab_c, list_sort_c) # Receptive Vocabulary * Working Memory
effect_plot(mods2, card_sort_c, pattern_c) # Cognitive Flexibility * Processing Speed
effect_plot(mods2, pattern_c, card_sort_c) # Processing Speed * Cognitive Flexibility
effect_plot(mods2, flanker_c, best_win) # Inhibitory Control * Words in Noise Threshold
effect_plot(mods2, best_win, flanker_c) # Words in Noise Threshold * Inhibitory Control
effect_plot(mods2, vocab_c, pattern_c) # Receptive Vocabulary * Processing Speed
effect_plot(mods2, pattern_c, vocab_c) # Processing Speed * Receptive Vocabulary
effect_plot(mods2, list_sort_c, card_sort_c) # Working Memory * Cognitive Flexibility
effect_plot(mods2, card_sort_c, list_sort_c) # Cognitive Flexibility * Working Memory

```

```{r, eval = F, warning = F, message = F}

# Combining initial model figures and saving
p1 / p2
ggsave("Plots/Initial_mods.png", plot = last_plot(), width = 8, height = 10, units = "in")

```

```{r, echo = F, warning = F, message = F}

# Merging initial model summaries together
initial_mods <- rbind(AM1_initial_mod, PDM10_initial_mod)

# Removing unneeded items from Environment
rm(p1, p2, mods1, mods2, AM1_initial_mod, PDM10_initial_mod)

```

## Models of Cognition and Intelligibility Change

### AM1 Models

Two elastic net models were used to model the relationship between the listener cognitive measures and intelligibility improvement (controlling for pretest intelligibility) for AM1: the main effects only model and the model with interactions. The difference in RMSE and R-squared is not significantly different. However, multiple interactions were indicated to be important predictors in the model.

The relative importance of the model variables are shown in the figure below. The most important variable in the model is the interaction between inhibitory control and words in noise hearing threshold followed by the interaction between processing speed and words in noise threshold, the interaction between cognitive flexibility and words in noise hearing threshold, and the interaction between inhibitory control and receptive vocabulary.

```{r, warning = F, message = F}

mod_fun2 <- function(speaker, title) {
  subsetted = cor_df %>%
    dplyr::select(-c(fluid_cog_c, pic_seq_c, improvement)) %>%
    dplyr::mutate(pwc_diff = resid(lm(posttest ~ pretest))) %>%
    dplyr::select(pwc_diff, ends_with("_c"), speaker, best_win) %>%
    tidyr::drop_na(pwc_diff, ends_with("_c"), speaker, best_win) %>%
    dplyr::filter(speaker == {{speaker}}) %>%
    dplyr::select(-speaker) %>%
    rename(pwc = pwc_diff)

  mod.1 = training(subsetted, "glmnet", int = F)
  mod.2 = training(subsetted, "glmnet", int = T)

  return(list(mod.1, mod.2))

}

mods12 = mod_fun2("AM1")
p12 = plot_fun(mods12, "a) Ataxic")
p12

```

```{r warning = F, message = F}

AM1_change_mod <- in_and_out(mods12, "Ataxic")

AM1_change_mod %>%
  gt::gt() %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE")

```

```{r, warning = F, message = F}

t_for_diff(mods12)

```

The the interaction effects for the intelligibility improvement model for AM1 are plotted below.

-   The interaction between inhibitory control and words in noise threshold indicate that the effect of inhibiotry control on intelligibility improvement was reduced for listeners with higher hearing thresholds
-   The interaction between processing speed and words in noise threshold indicate that greater processing speed reduced the effect on intelligibility improvement for listeners with higher thresholds.***
- The interaction between cognitive flexibility and words in noise threshold indicate that greater cognitive flexibility helped compensate for intelligibility improvement for listeners with high thresholds. However, for listeners with lower thresholds, greater cognitive flexibility had a negative effect on improvement.
-   The interaction between inhibitory control and receptive vocabulary indicates that increased vocabulary knowledge had a positive effect on intelligibility improvement for listeners with low and high inhibitory control. However, vocabulary knowledge had a negative effect for listeners with "mid/average" inhibitory control.

```{r, warning = F, message = F}

## AM1 Effect Plots

effect_plot(mods12, flanker_c, best_win) # Inhibitory Control * Words in Noise Threshold
effect_plot(mods12, best_win, flanker_c) # Words in Noise * Inhibitory Control
effect_plot(mods12, pattern_c, best_win, num = 4) # Processing Speed * Words in Noise Threshold
effect_plot(mods12, best_win, pattern_c, num = 4)
effect_plot(mods12, card_sort_c, best_win, num = 5) # Cognitive Flexibility * Words in Noise Threshold
effect_plot(mods12, flanker_c, vocab_c, num = 5) # Inhibitory Control * Receptive Vocabulary
effect_plot(mods12, vocab_c, flanker_c, num = 5) # Receptive Vocabulary * Inhibitory Control

```

### PDM10 Models

Two elastic net models were used to model the relationships between listener cognition measures and intelligibility improvement (controlling for pretest intelligibility) for PDM10. There is not a significant difference between the R-squared values for the two models.

The relative importance of the model variables is shown in the figure below. The most important variable in the model is the interaction between cognitive flexibility and receptive vocabulary followed by the interaction between inhibitory control and processing speed, the main effect of receptive vocabulary, the main effect of inhibitory control, and the interaction between inhibitory control and words in noise hearing threshold.

```{r, warning = F, message = F}

mods22 = mod_fun2("PDM10")
p22 = plot_fun(mods22, "b) PD")
p22

```

```{r, warning = F, message = F}

PDM10_change_mod <- in_and_out(mods22, "PD")

PDM10_change_mod %>%
  gt::gt() %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE")

```

```{r, warning = F, message = F}

t_for_diff(mods22)

```

The interaction effects for the intelligibility improvement models for PDM10 are plotted below.

-   The interaction between cognitive flexibility and receptive vocabulary indicates that cognitive flexibility had a positive effect on improvement for listeners with higher vocabulary knowledge, but minimal effect for listeners with lower vocabulary knowledge
-   The interaction effect between inhibitory control and processing speed indicates that processing speed has a negative effect on intelligibility improvement for listeners with mid-range inhibitory control. However, processing speed has minimal effect on improvement when listeners have lower or higher inhibitory control scores.***
-   Receptive vocabulary has a positive effect on intelligibility improvement.
-   Inhibitory control has a negative effect on intelligibility improvement.
-   The interaction between inhibitory control and words in noise threshold indicates a positive effect of inhibitory control for listeners with low words in noise thresholds but little to no effect for listeners with elevated hearing thresholds (a potential negative effect)

```{r, warning = F, message = F}

## PDM10 Effect Plots

effect_plot(mods22, card_sort_c, vocab_c) # Cognitive Flexibility * Receptive Vocabulary
effect_plot(mods22, vocab_c, card_sort_c) # Receptive Vocabulary * Cognitive Flexibility
effect_plot(mods22, flanker_c, pattern_c) # Inhibitory Control * Processing Speed
effect_plot(mods22, pattern_c, flanker_c) # Processing Speed * Inhibitory Control
effect_plot(mods22, vocab_c, int = F) # Receptive Vocab 
effect_plot(mods22, flanker_c, int = F) # Inhibitory Control
effect_plot(mods22, flanker_c, best_win, num = 4) # Inhibitory Control * Words in Noise Threshold

```

```{r, eval = F, warning = F, message = F}

# Combining and saving change model figures
p12 / p22
ggsave("Plots/Change_mods.png", plot = last_plot(), width = 8, height = 10, units = "in")

```

```{r, echo = F, warning = F, message = F}

# Merging Model summaries together
change_mods <- rbind(AM1_change_mod, PDM10_change_mod)

# Creating Overall Model Table

initial_mods <- initial_mods %>%
  dplyr::mutate(outcome = "Initial Intelligibility")

change_mods <- change_mods %>%
  dplyr::mutate(outcome = "Intelligbility Change")

mods_table <- rbind(initial_mods, change_mods)

mods_table %>%
  gt::gt(groupname_col = "outcome") %>%
  cols_label(speaker = "Speaker",
             mods = "Model Specification",
             out_sample_r2 = "R\U00B2",
             out_sample_rmse = "RMSE") %>%
  fmt_number(decimals = 2) %>%
  gt::gtsave("Tables/mods_table.html")

# Removing unneeded items from environment
rm(p12, p22, AM1_change_mod, PDM10_change_mod, mods12, mods22)

```